02 - PubliliaFitness - Defining strategies and calculating frequencies

### 2023-08-25 scripts/Publilia-02-strats.Rmd
minor edits to ensure scripts 01-03 run after overhaul of strats_data

---
title: '01 - TreehopperFitness: Defining strategies and calculating frequencies'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

@@

## Required libraries and functions
```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggh4x)
library(rstatix)
library(emmeans)
library(effectsize)
library(gt)
library(gtsummary)
theme_gtsummary_compact()
library(flextable)
library(multcompView)
library(here)

@@

clutch_status_guarding_summary_v2
```{r}
All_data_v6 <- read_rds(here("saved_rds/All_data_v6.rds"))
# clutch_status_guarding_summary_v2 <- read_rds(here("saved_rds/clutch_status_guarding_summary_v2.rds"))
# last_day_fID <- read_rds(here("saved_rds/last_day_fID.rds"))

strats_data <- read_rds(here("saved_rds/strats_data.rds"))
```

@@

# 3 - Defining exposure to ants

## 3.1 - Ant counts on plant; some ants_removed clutches were visited by ants, and vice versa
```{r}
## Moved to Publilia-01-qc.Rmd
# ant_counts <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch) %>%
#     summarize(ants_removed=first(ants_removed),
#               females_removed=first(females_removed),
#               n_antsP=sum(ants_P,na.rm=T),
#               ever_antsP=n_antsP>0,
#               mean_antsP=mean(ants_P,na.rm=T),
#               med_antsP=median(ants_P,na.rm=T),
#               
#               n_antsP_b4h=sum(ants_P&b4h),
#               ever_antsP_b4h=n_antsP_b4h>0,
#               mean_antsP_b4h=mean(ifelse(b4h,ants_P,NA),na.rm=T),
#               med_antsP_b4h=median(ifelse(b4h,ants_P,NA),na.rm=T),
#               
#               n_antsP_afh=sum(ants_P&afh),
#               ever_antsP_afh=n_antsP_afh>0,
#               mean_antsP_afh=mean(ifelse(afh,ants_P,NA),na.rm=T),
#               med_antsP_afh=median(ifelse(afh,ants_P,NA),na.rm=T)) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

ggplot(data=ant_counts,aes(x=Dorm,y=n_antsP, fill=ants_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP, fill=ants_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP_b4h, fill=ants_treatment)) +
  geom_boxplot()

ggplot(data=ant_counts,aes(x=Dorm,y=mean_antsP_afh, fill=ants_treatment)) +
  geom_boxplot()

ant_counts %>%
  count(Dorm,ants_treatment,ever_antsP)

ant_counts %>%
  count(Dorm,ants_treatment,ever_antsP_b4h)

ant_counts %>%
  count(Dorm,ants_treatment,ever_antsP_afh)

# how many clutches ever had ants in each of the dorms?
ant_counts %>%
  ggplot(data=., aes(y=as.numeric(ever_antsP), x=Dorm, fill=ants_treatment)) +
  stat_summary(fun="mean", geom = "bar", position = position_dodge(0.9)) +
  stat_summary(fun.data="mean_se", geom = "errorbar",position = position_dodge(0.9), width = 0.2) +
  stat_summary(fun.data="n", geom = "text", position = position_dodge(0.9)) +
  theme_publilia()

ant_counts %>%
  ggplot(data=., aes(y=as.numeric(ever_antsP_b4h), x=Dorm, fill=ants_treatment)) +
  stat_summary(fun="mean", geom = "bar", position = position_dodge(0.9)) +
  stat_summary(fun.data="mean_se", geom = "errorbar",position = position_dodge(0.9), width = 0.2) +
  stat_summary(fun.data="n", geom = "text", position = position_dodge(0.9)) +
  theme_publilia()

ant_counts %>%
  ggplot(data=., aes(y=as.numeric(ever_antsP_afh), x=Dorm, fill=ants_treatment)) +
  stat_summary(fun="mean", geom = "bar", position = position_dodge(0.9)) +
  stat_summary(fun.data="mean_se", geom = "errorbar",position = position_dodge(0.9), width = 0.2) +
  stat_summary(fun.data="n", geom = "text", position = position_dodge(0.9)) +
  theme_publilia()

# how extreme was the break-through?
ant_counts %>%
  ggplot(data=., aes(x=mean_antsP_b4h, color=ants_treatment, linetype=females_removed, group=Dorm)) +
  geom_freqpoly(bins=100)

ant_counts %>%
  ggplot(data=., aes(x=med_antsP_b4h, color=ants_treatment, linetype=females_removed, group=Dorm)) +
  geom_freqpoly()

ant_counts %>%
  ggplot(data=., aes(x=mean_antsP_afh, color=ants_treatment, linetype=females_removed, group=Dorm)) +
  geom_freqpoly(bins=100)

ant_counts %>%
  ggplot(data=., aes(x=med_antsP_afh, color=ants_treatment, linetype=females_removed, group=Dorm)) +
  geom_freqpoly()

# how many clutches actually ever had ants vs. not?
ant_counts %>%
  ggplot(data=., aes(x=ever_antsP_b4h, fill=ants_treatment)) +
  geom_bar() +
  theme_publilia() + facet_wrap(~females_removed)

ant_counts %>%
  ggplot(data=., aes(x=ever_antsP_afh, fill=ants_treatment)) +
  geom_bar() +
  theme_publilia() + facet_wrap(~females_removed)

ant_counts %>%
  ggplot(data=., aes(x=ever_antsP_b4h|ever_antsP_afh, fill=ants_treatment)) +
  geom_bar() +
  theme_publilia() + facet_wrap(~females_removed)

## 3.2 - Ant arrival times
```{r}
## Moved to Publilia-01-qc.Rmd
# antP_arrival_afh_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsP_afh=ants_P&afh) %>%
#     filter(is_antsP_afh) %>%
#     slice_head(n=1) %>% 
#   mutate(antP_arrival_afh=Date-est_hatch) %>%
#   select(location, antP_arrival_afh, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antP_last_b4h_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsP_bfh=ants_P&b4h) %>%
#     filter(is_antsP_bfh) %>%
#     slice_tail(n=1) %>% 
#   mutate(antP_last_b4h=Date-est_hatch) %>%
#   select(location, antP_last_b4h, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antL_arrival_afh_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsL_afh=ants_L&afh) %>%
#     filter(is_antsL_afh) %>%
#     slice_head(n=1) %>% 
#   mutate(antL_arrival_afh=Date-est_hatch) %>%
#   select(location, antL_arrival_afh, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# antL_last_b4h_data <- All_data_v6 %>%
#   left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
#   filter(!is.na(initiation_date)) %>%
#   distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
#   group_by(location) %>%
#     mutate(b4h=Date>=initiation_date&Date<est_hatch,
#            afh=Date>=est_hatch,
#            is_antsL_bfh=ants_L&b4h) %>%
#     filter(is_antsL_bfh) %>%
#     slice_tail(n=1) %>% 
#   mutate(antL_last_b4h=Date-est_hatch) %>%
#   select(location, antL_last_b4h, ants_removed, females_removed,
#          initiation_date, est_hatch) %>%
#   ungroup() %>%
#   mutate(Dorm=substr(location,1,1),
#          ants_treatment=ifelse(ants_removed,"Ants removed","Control"))
# 
# ant_data <- ant_counts %>%
#   left_join(antP_last_b4h_data) %>%
#   left_join(antP_arrival_afh_data) %>%
#   left_join(antL_last_b4h_data) %>%
#   left_join(antL_arrival_afh_data)
# 
# visited_1st_4d_data_ants <- visited_1st_4d_data %>%
#   left_join(ant_data)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_last_b4h, color=females_removed)) +
  geom_point() +
  geom_abline(slope=1)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h, y=antL_arrival_afh, color=females_removed)) +
  geom_point() +
  geom_smooth(method=lm) +
  geom_abline(slope=1)

strats_data %>%
  filter(ever_antsP) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_last_b4h)) +
  geom_boxplot()

strats_data %>%
  filter(ants=="Control") %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)

strats_data %>%
  filter(ants=="Control", !females_removed) %>%
  ggplot(data=., aes(x=est_rel2h>=0, y=antL_arrival_afh)) +
  geom_boxplot() +
  geom_jitter(pch = 21, size=1, width=0.35, height=0) +
  stat_summary(fun=mean, geom="point", shape=4, size=3)
```

@@

#### Desertion relative to hatch
Using geom_area and density
```{r}
Fig1_v1_data <- visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

strat_data_table1 <- visited_1st_4d_data %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Females not removed (A)` = !females_removed,
         `Leaf hosted a clutch (B)` = !is.na(initiation_date),
         `Neither (!A&!B)` = `Females not removed (A)`&`Leaf hosted a clutch (B)`) %>%
  select(`Ant exclusion`, `Females not removed (A)`, `Leaf hosted a clutch (B)`,
         `Neither (!A&!B)`) %>%
  tbl_summary(by=`Ant exclusion`)

@@

strat_data_table2 <- visited_1st_4d_data %>%
  filter(!is.na(initiation_date), !females_removed) %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Mother at initiation (A)` = !is.na(clutch_initiated_order),
         `First clutch initiated (B)`=clutch_initiated_order==1,
         `Didn't disappear (C)`=visit_ends<last_day_fID,
         `Marked (D)`=!grepl("U", Female_ID),
         `Visited early (E)`=visited_1st_4d,
         `Figure 3A (A&B&C&D&!E)`=`Mother at initiation (A)`&`First clutch initiated (B)`&
           `Didn't disappear (C)`&`Marked (D)`&!`Visited early (E)`,
         `Figure 3B (A&B&C&D&E)`=`Mother at initiation (A)`&`First clutch initiated (B)`&
           `Didn't disappear (C)`&`Marked (D)`&`Visited early (E)`,
         `Figure S4C (!A&C)` = !`Mother at initiation (A)`&`Didn't disappear (C)`) %>%
  select(`Ant exclusion`, `Mother at initiation (A)`, `First clutch initiated (B)`,
         `Didn't disappear (C)`, `Marked (D)`,`Visited early (E)`, 
         `Figure 3A (A&B&C&D&!E)`, `Figure 3B (A&B&C&D&E)`, `Figure S4C (!A&C)`) %>%
  tbl_summary(by=`Ant exclusion`, missing="no")

@@

Fig1_v1C.gd <- ggplot(data = filter(visited_1st_4d_data,
                                 !females_removed, !mom_at_init,
                                 visit_ends >= initiation_date,
                                 visit_ends < last_day_fID)) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof secondary visitors") +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

@@

filter(clutch_status_guarding_summary_v2, location == "A18_8", mom_at_init) %>%
  select(Female_ID:hatch_date, first_visitor_initiator, init_dur_rank:init_first_dur_shortest)
```

@@

### Non-parametric stats: v1
```{r}
# non-parametric: Kruskall-Wallis test followed by Dunn test or pairwise wilcox tests
vis_cat_v1.data <- visited_1st_4d_data %>%
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  filter(!females_removed, mom_at_init, 
         clutch_initiated_order == 1, 
         !grepl("U", Female_ID),
         visit_ends < last_day_fID,
         # ** There was one female (E_Gkgg) which was on the eggs before and after but not ON the initiation date.
         # I suspect I missed her in the field or she may have left and come back. In any case she was never visited
         # and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.
         !is.na(vis_cat))

@@

### Non-parametric stats, but including females that disappeared or were unmarked
Three clutches were omitted from the analysis. Two were visited after hatching, and one had no females present at initiation.
```{r}
# non-parametric: Kruskall-Wallis test followed by Dunn test or pairwise wilcox tests
vis_cat_v2.data <- visited_1st_4d_data %>%
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = ifelse(location=='E40_9', "never visited", vis_cat),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  filter(!females_removed, mom_at_init, 
         clutch_initiated_order == 1, !is.na(vis_cat))

# # where do the NAs come from?
# vis_cat_v2.data %>% filter(is.na(vis_cat))
# visited_1st_4d_data %>% filter(location=='C41_6')
# # C41_6 was deserted immediately, then visited long after hatch
# visited_1st_4d_data %>% filter(location=='E148_6')
# # E148_6 was also visited long after hatch
# visited_1st_4d_data %>% filter(location=='E40_9')
# # There was one female (E_Gkgg: E40_9) on the eggs before and after but not ON the init date.
# # I suspect I missed her in the field or she may have left and come back. She was never visited
# # and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.

@@

# 6 Comparing early desertion by initiator vs. latest desertion of leaf by any female vs. latest desertion of the plant by any female
```{r}
visit_data.desL.desP <- visited_1st_4d_data %>%
  left_join(first_visit_cuts) %>%
  group_by(location) %>% mutate(last_visit_ends_L = max(visit_ends, na.rm=T)) %>% ungroup() %>%
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  group_by(DormPlant) %>% mutate(last_visit_ends_P = max(visit_ends, na.rm=T)) %>% ungroup() %>%
  mutate(last_desL_rel2h = last_visit_ends_L-est_hatch,
         last_desP_rel2h = last_visit_ends_P-est_hatch)

@@

## Original from chapter:
```{r, fig.width=3, fig.height=3}
Fig1_v1_data <- visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

strat_data_table1 <- visited_1st_4d_data %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Females not removed (A)` = !females_removed,
         `Leaf hosted a clutch (B)` = !is.na(initiation_date),
         `Neither (!A&!B)` = `Females not removed (A)`&`Leaf hosted a clutch (B)`) %>%
  select(`Ant exclusion`, `Females not removed (A)`, `Leaf hosted a clutch (B)`,
         `Neither (!A&!B)`) %>%
  tbl_summary(by=`Ant exclusion`)

@@

visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited within first 4 days", 
                  "Not visited within first 4 days")) %>%
  ggplot(aes(x=est_rel2h, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Number of females") +
  scale_y_continuous(breaks = c(0,4,8),
                     labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        legend.justification = c(1,1),
        legend.position = c(1,0.2)) +
  geom_text(data = init_text,
            aes(x = -20, y = 10, label = label),
            hjust = 0, vjust = -1, color = "gray", fontface = "bold") +
  geom_text(data = hatch_text,
            aes(x = 0, y = 10, label = label),
            hjust = 0, vjust = -1, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off")
```

@@

(guarding_durations_plot_2023 <- visited_1st_4d_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited\nwithin first 4 days", 
                  "Not visited\nwithin first 4 days")) %>%
  ggplot(aes(x=guard_duration, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Number of females") +
  scale_y_continuous(breaks = c(0,4,8),
                     labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        strip.text = element_text(size = 12),
        legend.justification = c(1,1),
        legend.position = c(1,0.2),
        legend.text = element_text(size = 12)) +
  geom_text(data = hatch_text,
            aes(x = 20, y = Inf, label = label),
            hjust = -0.1, vjust = 0, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off"))
```

>>>>

---
title: '02 - PubliliaFitness - Defining strategies and calculating frequencies'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

@@

## Required libraries and functions
```{r}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(ggh4x)
library(rstatix)
library(emmeans)
library(effectsize)
library(gt)
library(gtsummary)
theme_gtsummary_compact()
library(flextable)
library(multcompView)
library(scales)
library(ggupset)
library(here)

@@

All_data_v6 and strats_data should be all we need
```{r}
All_data_v6 <- read_rds(here("saved_rds/All_data_v6.rds"))
strats_data <- read_rds(here("saved_rds/strats_data.rds"))
```

@@

@@

#### Desertion relative to hatch
Using geom_area and density
```{r}
Fig1_v1_data <- strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

strat_data_table1 <- strats_data %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Females not removed (A)` = !females_removed,
         `Leaf hosted a clutch (B)` = !is.na(initiation_date),
         `Neither (!A&!B)` = `Females not removed (A)`&`Leaf hosted a clutch (B)`) %>%
  select(`Ant exclusion`, `Females not removed (A)`, `Leaf hosted a clutch (B)`,
         `Neither (!A&!B)`) %>%
  tbl_summary(by=`Ant exclusion`)

@@

strat_data_table2 <- strats_data %>%
  filter(!is.na(initiation_date), !females_removed) %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Mother at initiation (A)` = !is.na(clutch_initiated_order),
         `First clutch initiated (B)`=clutch_initiated_order==1,
         `Didn't disappear (C)`=visit_ends<last_day_fID,
         `Marked (D)`=!grepl("U", Female_ID),
         `Visited early (E)`=visited_1st_4d,
         `Figure 3A (A&B&C&D&!E)`=`Mother at initiation (A)`&`First clutch initiated (B)`&
           `Didn't disappear (C)`&`Marked (D)`&!`Visited early (E)`,
         `Figure 3B (A&B&C&D&E)`=`Mother at initiation (A)`&`First clutch initiated (B)`&
           `Didn't disappear (C)`&`Marked (D)`&`Visited early (E)`,
         `Figure S4C (!A&C)` = !`Mother at initiation (A)`&`Didn't disappear (C)`) %>%
  select(`Ant exclusion`, `Mother at initiation (A)`, `First clutch initiated (B)`,
         `Didn't disappear (C)`, `Marked (D)`,`Visited early (E)`, 
         `Figure 3A (A&B&C&D&!E)`, `Figure 3B (A&B&C&D&E)`, `Figure S4C (!A&C)`) %>%
  tbl_summary(by=`Ant exclusion`, missing="no")

@@

Fig1_v1C.gd <- ggplot(data = filter(strats_data,
                                 !females_removed, !mom_at_init,
                                 visit_ends >= initiation_date,
                                 visit_ends < last_day_fID)) +
  geom_area(aes(x = guard_duration, y=..density.., color = ants, fill = ants), 
            alpha = 0.75, stat='bin', binwidth = 4, size = 0, position = "identity") + 
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Proportion\nof secondary visitors") +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1)

@@

filter(strats_data, location == "A18_8", mom_at_init) %>%
  select(Female_ID:hatch_date, first_visitor_initiator, init_dur_rank:init_first_dur_shortest)
```

@@

### Non-parametric stats: v1
```{r}
# non-parametric: Kruskall-Wallis test followed by Dunn test or pairwise wilcox tests
vis_cat_v1.data <- strats_data %>%
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  filter(!females_removed, mom_at_init, 
         clutch_initiated_order == 1, 
         !grepl("U", Female_ID),
         visit_ends < last_day_fID,
         # ** There was one female (E_Gkgg) which was on the eggs before and after but not ON the initiation date.
         # I suspect I missed her in the field or she may have left and come back. In any case she was never visited
         # and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.
         !is.na(vis_cat))

@@

### Non-parametric stats, but including females that disappeared or were unmarked
Three clutches were omitted from the analysis. Two were visited after hatching, and one had no females present at initiation.
```{r}
# non-parametric: Kruskall-Wallis test followed by Dunn test or pairwise wilcox tests
vis_cat_v2.data <- strats_data %>%
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = ifelse(location=='E40_9', "never visited", vis_cat),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  filter(!females_removed, mom_at_init, 
         clutch_initiated_order == 1, !is.na(vis_cat))

# # where do the NAs come from?
# vis_cat_v2.data %>% filter(is.na(vis_cat))
# strats_data %>% filter(location=='C41_6')
# # C41_6 was deserted immediately, then visited long after hatch
# strats_data %>% filter(location=='E148_6')
# # E148_6 was also visited long after hatch
# strats_data %>% filter(location=='E40_9')
# # There was one female (E_Gkgg: E40_9) on the eggs before and after but not ON the init date.
# # I suspect I missed her in the field or she may have left and come back. She was never visited
# # and deserted 20 days after initiation and we have plenty such data points, so I think it's safe to ignore.

@@

# 6 Comparing early desertion by initiator vs. latest desertion of leaf by any female vs. latest desertion of the plant by any female
```{r}
visit_data.desL.desP <- strats_data %>%
  group_by(location) %>% mutate(last_visit_ends_L = max(visit_ends, na.rm=T)) %>% ungroup() %>%
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  group_by(DormPlant) %>% mutate(last_visit_ends_P = max(visit_ends, na.rm=T)) %>% ungroup() %>%
  mutate(last_desL_rel2h = last_visit_ends_L-est_hatch,
         last_desP_rel2h = last_visit_ends_P-est_hatch)

@@

## Original from chapter:
```{r, fig.width=3, fig.height=3}
Fig1_v1_data <- strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID)

strat_data_table1 <- strats_data %>%
  mutate(`Ant exclusion` = ifelse(ants_removed, "Ants excluded", "Control"),
         `Females not removed (A)` = !females_removed,
         `Leaf hosted a clutch (B)` = !is.na(initiation_date),
         `Neither (!A&!B)` = `Females not removed (A)`&`Leaf hosted a clutch (B)`) %>%
  select(`Ant exclusion`, `Females not removed (A)`, `Leaf hosted a clutch (B)`,
         `Neither (!A&!B)`) %>%
  tbl_summary(by=`Ant exclusion`)

@@

strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited within first 4 days", 
                  "Not visited within first 4 days")) %>%
  ggplot(aes(x=est_rel2h, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Desert date relative to hatching") +
  ylab("Number of females") +
  scale_y_continuous(breaks = c(0,4,8),
                     labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = -20, linetype="dotted", 
                color = "gray", size=1) +
  geom_vline(xintercept = 0, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        legend.justification = c(1,1),
        legend.position = c(1,0.2)) +
  geom_text(data = init_text,
            aes(x = -20, y = 10, label = label),
            hjust = 0, vjust = -1, color = "gray", fontface = "bold") +
  geom_text(data = hatch_text,
            aes(x = 0, y = 10, label = label),
            hjust = 0, vjust = -1, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off")
```

@@

(guarding_durations_plot_2023 <- strats_data %>%
  filter(!grepl("U", Female_ID), !females_removed, mom_at_init,
         clutch_initiated_order == 1, visit_ends < last_day_fID) %>% 
  filter(!is.na(`visited in 4 days`)) %>%
  mutate(`visited in 4 days` = 
           ifelse(`visited in 4 days` == "Visited", 
                  "Visited\nwithin first 4 days", 
                  "Not visited\nwithin first 4 days")) %>%
  ggplot(aes(x=guard_duration, fill=ants)) +
  geom_histogram(binwidth = 4) +
  theme_publilia() +
  xlab("Guarding duration") +
  ylab("Number of females") +
  scale_y_continuous(breaks = c(0,4,8),
                     labels = label_number(accuracy = 1)) +
  scale_color_manual(values = c("black", "brown1")) +
  scale_fill_manual(values = c("black", "brown1")) +
  geom_vline(xintercept = 20, linetype="dashed", 
                color = "gray", size=1) +
  ggh4x::facet_grid2(vars(`visited in 4 days`, ants), 
                     strip = hist_strips) +
  theme(strip.background = element_rect(color = NA),
        strip.text = element_text(size = 12),
        legend.justification = c(1,1),
        legend.position = c(1,0.2),
        legend.text = element_text(size = 12)) +
  geom_text(data = hatch_text,
            aes(x = 20, y = Inf, label = label),
            hjust = -0.1, vjust = 0, color = "gray", fontface = "bold") +
  coord_cartesian(clip = "off"))
```
